|

### 2021-10-02 appendix/appendix.Rmd
small edits to appendix text

The expected utility for the city councilors depends on which reversal policy takes place. As mentioned above, we explore the implications of three different mechanisms. First, in the \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminishing by a factor $\delta \in (0,1)$ and with the random selection of one councilor. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If rejected, the budget shrinks again by a factor $\delta$, and another city councilor is chosen to propose the policy. The process continues until one proposal is finally accepted.

@@

We derive the optimal rents ($r$) and public goods provision ($g$) a mayor should propose. Mayors benefits from public goods provision, as it increases their chances of reelection. However, they would rather provide the lowest necessary amount of public goods and extract the remaining resources as political rents. As we noted previously, their objective is to maximize their expected utility, subject to the municipal budget constraint.

@@

\begin{proposition}
Public goods provision increases with legislature size if governing costs decrease with legislature size.
\end{proposition}

@@

We place the offer on the left-hand side. There are two components on the right-hand side. The first is the amount that the councilor $i$ receives when he/she is the proposer. The amount is equal to the chance that they are recognized as the proposer, times the budget in the next round, minus the offers they have to make to convince half of the councilors. The second part denotes the gains obtained by the councilor if he/she rejects the current offer but still receives an offer in the following round. This is equal to the chance that they are not recognized as the proposer, times the chance that they receive a transfer, times the transfer amount. Note that the councilor's offer as the proposer is the same as the offer they want to receive, as city councilors are exchangeable. The solution is symmetric for all councilors receiving an offer (this means that, without loss of generality, we could have dropped the $i$ in the solution). After some calculations, the offer $x_i$ has to be greater than or equal to:

@@

We use three data sources in our work, mainly from Brazilian governmental agencies. Table \ref{tab:varsources} displays the primary sources we use here and their respective URLs.

@@

\begin{itemize}
\item \textbf{Number of Seats 2000}: Number of city councilors in 2000 (source: TSE).
\item \textbf{Population 2000}: City population according to the 2000 Brazilian Census (source: IBGE).
\item \textbf{Per-Capita GDP 2000 (in millions)}: Municipal per capita GDP measured by the Brazilian Census (source: IPEA).
\item \textbf{Proportion of Poverty 2000}: Percentage of citizens earning less than R\$ 70.00 per month (source: Social Development Ministry [MDS]).
\item \textbf{Number of Seats 2004}: Number of city councilors in 2004 according to the Electoral Court (source: TSE).
\item \textbf{Infant Mortality 2005--2008}: Number of children who died before reaching one year of age, divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
\item \textbf{Postneonatal Mortality 2005--2008}: Number of children alive for at least 28 days who died before reaching one year of age, divided by the total number of children who lived for 28 days, multiplied by 1,000 (source: DataSUS)
\item \textbf{Enrollment Elementary School 2005--2008}: Number of children enrolled in elementary schools, averaged by classroom size (K--4) (source: INEP).
\item \textbf{Quality of Elementary School 2005--2008}: Average municipal IDEB scores. IDEB scores are estimated as follows. The Ministry of Education takes the average of the students' grades in math and Portuguese, then multiply the those numbers by the student yearly approval rate. INEP reweights the estimators to avoid schools and classroom specific effects (source: INEP, years: 2005, 2007).
\item \textbf{Size of Mayoral Pre-Electoral Coalition 2004}: Number of elected councilors in the mayoral pre-electoral coalition (source: TSE)
\item \textbf{Number of Appointed Bureaucrats 2005--2008}: Number of bureaucrats of the direct administration who were appointed to jobs in the municipality (source: IBGE, years: 2005, 2006, and 2008)
\item \textbf{Number of Females Elected 2004}: Number of females elected to city council (source: TSE)
\item \textbf{Number of Non-Whites Elected 2004}: Number of non-whites elected to the city council. We collect this data for municipalities less than 10,000 inhabitants away from the cutoffs (source: own compilation based on the TSE candidate pictures).
\item \textbf{Competition per Seat 2004}: Number of people running for city councilor divided by the city council size (source: TSE)
\item \textbf{Proportion Approved Legislation 2005}: Number of approved legislation in 2005 divided by the number of proposed legislation (source: InterLegis)
\end{itemize}

@@

\begin{itemize}
\item \textbf{Councilors Job Appointments}: How often mayors appoint people allied to city councilors to bureaucratic jobs (source: survey).
\item \textbf{Councilors Demands}: How often mayors satisfy an allied councilors' demand (e.g., build a health clinic in a councilor's neighborhood) (source: survey).
\item \textbf{Councilors Personal Requests}: How often mayors satisfy an allied councilors' requests (e.g., hire an ally in the bureaucracy) (source: survey).
\item \textbf{Councilors Legislation}: How often that mayors support the councilors' legislative agenda (source: survey).
\item \textbf{Councilors Constructions}: How often mayors support councilors' requests for constructions (e.g., fix potholes in the councilor's neighborhoods) (source: survey).
\end{itemize}

This Appendix contains other questions that we asked in the survey. They are meant to improve our knowledge about the city councilors' perceptions of the roles the municipality and themselves play in the government. We omitted these questions in the main paper because they are not of general interest to our argument. The descriptive statistics of the variables used in the main article follow in Table \ref{tab:descrvarspaper}.

@@

For the first assumption, \citet{McCrary2008} proposes a measure of the distributional imbalance around the discontinuity, testing whether cases are more abundant on either side of the cutoff. In our sample, the McCrary statistic is 0.391 (SE $=$ 0.299), showing no evidence of manipulation.

We run the \citet{cattaneo2019simple} manipulation test, which is based on the density of the local polynomial estimator. We use local polynomial orders from one to four. As the null hypothesis is no manipulation, the p-values for each polynomial order are: local linear (p-value $=0.442$); quadratic (p-value $=0.740$); cubic (p-value $=0.998$); and quartic (p-value $=0.620$). Therefore, we also see no evidence of manipulation.

@@

As Brazil has only a few municipalities with a population higher than one million inhabitants, we focus on cities with less than $571,429 + 47,619 = 619,048$ inhabitants. Table \ref{tab:citycouncilsize2004short} displays the number of municipalities by the thresholds we use in this paper.

@@

To provide an overview of the data dispersion, Figure \ref{sampleplot} plots the municipalities by their proximity to the cutoffs. In the map, we give the contour of every Brazilian city, coloring it by the proximity to the council size thresholds. Darker colors represent municipalities closer to the population thresholds. In regression discontinuity models, the closer a city is to the cutoff, the more influential it is in the estimation. Figure \ref{sampleplot} also shows that the municipalities around the cutoff are reasonably distributed across the country.

@@

When estimating regression discontinuities with multiple cutoffs, authors pool all the discontinuities together and estimate the effect of council size on outcomes \citep{cattaneo2016interpreting}. Let $C = \{c_0, c_1, \cdots, c_n, c_{n+1}\}$ be the set of cutoffs thresholds associated with the running variable $X$. The outcome variable is assumed to be $Y$. In the pooled model, the Local Average Treatment Effect (LATE) is:

@@

Nevertheless, this may be problematic as municipal characteristics may vary between cutoffs. For instance, population imbalance is a known source of heterogeneity, as shown by \citet{Cattaneo2016}. To ensure that cases are comparable, we estimate the average LATE for all cutoffs. We add a set of controls $Z_i$ to reduce heterogeneity. For instance, the population that determines the change from one threshold to another also determines all the cutoffs. As the population runs smoothly around each cutoff, it makes this variable an ideal control to add to our models. The new estimator is:

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1\linewidth}
      \centerline{\includegraphics[width=5in,angle=0]{../figures/bwSens.pdf}}
      \caption{Bandwidth Sensitivity -- Main Models}
      \bigskip
      \label{graphbw}
    \end{minipage}
  \end{center}
\end{figure}

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[width=5in,angle=0]{../figures/placBwSens.pdf}}
      \caption{Bandwidth Sensitivity -- Placebo Regressions}
      \bigskip
      \label{placgraphbw}
    \end{minipage}
  \end{center}
\end{figure}

@@

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[height=8in,angle=0]{../figures/dropStateSens1.pdf}}
      \caption{Sensitivity Analysis for the States in the Sample -- Welfare Outcomes}
      \bigskip
      \label{graphsensstatedrop1}
    \end{minipage}
  \end{center}
\end{figure}

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[height=8in,angle=0]{../figures/dropStateSens2.pdf}}
      \caption{Sensitivity Analysis of State Characteristics -- Mechanism Outcomes}
      \bigskip
      \label{graphsensstatedrop2}
    \end{minipage}
  \end{center}
\end{figure}

\begin{figure}[htb!]
  \begin{center}
      \centerline{\includegraphics[width=5in,angle=0]{../figures/placDropStateSens1.pdf}}
      \caption{Sensitivity Analysis of State Characteristics with Placebo Cutoffs -- Welfare Outcomes}
      \bigskip
      \label{placgraphsensstatedrop1}
  \end{center}
\end{figure}

\begin{figure}[htb!]
  \begin{center}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/placDropStateSens2.pdf}}
      \caption{Sensitivity Analysis of State Characteristics with Placebo Cutoffs -- Mechanism Outcomes}
      \bigskip
      \label{placgraphsensstatedrop2}
  \end{center}
\end{figure}

@@

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[height=8in,angle=0]{../figures/ffSens.pdf}}
      \caption{Sensitivity Analysis of Model Functional Form}
      \bigskip
      \label{graphsensff}
    \end{minipage}
  \end{center}
\end{figure}

\begin{figure}[htb!]
  \begin{center}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/placFfSens.pdf}}
      \caption{Sensitivity Analysis of Model Functional Form -- Placebo Cutoffs}
      \bigskip
      \label{placgraphsensff}
  \end{center}
\end{figure}

@@

\subsection{Sensitivity Analysis of Additional Cutoffs\label{senscutapp}}

@@

@@

\begin{figure}[htb!]
\begin{center}
\centerline{\includegraphics[width=6in,angle=0]{../figures/rdAddCuts1.pdf}}
\caption{Sensitivity to Additional Cutoffs -- Welfare Outcomes}
\bigskip
\label{senscut1}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\centerline{\includegraphics[width=6in,angle=0]{../figures/rdAddCuts2.pdf}}
\caption{Sensitivity to Additional Cutoffs -- Mechanism Outcomes}
\bigskip
\label{senscut2}
\end{center}
\end{figure}

>>>>

The expected utility for the city councilors depends on which reversal policy takes place. As mentioned above, we explore the implications of three different mechanisms. First, in the \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminishing by a factor $\delta \in (0,1)$ and the random selection of one councilor. The randomly selected councilor becomes the new proposer. If her proposal is accepted, it is implemented. If rejected, the budget shrinks again by $\delta$ and another city councilor is chosen to propose the policy. The process continues until the proposal is finally accepted.

@@

We derive the optimal rents ($r$) and public goods provision ($g$) a mayor should propose. Mayors benefit from public goods provision, as it increases their chances of reelection. However, they would rather provide the lowest necessary amount of public goods and extract the remaining resources as political rents. As we noted previously, their objective is to maximize their expected utility, subject to the municipal budget constraint.

@@

\begin{proposition}
Public goods provision increases with larger legislatures if governing costs decrease in larger legislatures.
\end{proposition}

@@

We place the offer on the left-hand side. There are two components on the right-hand side. The first is the amount that councilor $i$ receives when he/she is the proposer. The amount is equal to the chance that they are recognized as the proposer, times the budget in the next round, minus the offers they have to make to convince half of the councilors. The second part denotes the gains obtained by the councilor if he/she rejects the current offer but still receives an offer in the following round. This is equal to the chance that they are not recognized as the proposer, times the chance that they receive a transfer, times the transfer amount. Note that the councilor's offer as the proposer is the same as the offer they want to receive, as city councilors are exchangeable. The solution is symmetric for all councilors receiving an offer (this means that, without loss of generality, we could have dropped the $i$ in the solution). After some calculations, the offer $x_i$ has to be greater than or equal to:

@@

We use three data sources in our work, mainly from Brazilian governmental agencies. Table \ref{tab:varsources} presents the variables we use here and their respective URLs.

@@

\begin{itemize}
\item \textbf{Number of Seats 2000}: Number of city councilors in 2000 (source: TSE).
\item \textbf{Population 2000}: City population according to the 2000 Brazilian Census (source: IBGE).
\item \textbf{Per Capita GDP 2000 (Millions)}: Municipal per capita GDP measured by the Brazilian Census (source: IPEA).
\item \textbf{Proportion of Poverty 2000}: Percentage of citizens earning less than R\$ 70.00 per month (source: Social Development Ministry [MDS]).
\item \textbf{Number of Seats 2004}: Number of city councilors in 2004 according to the Electoral Court (source: TSE).
\item \textbf{Infant Mortality 2005--2008}: Number of children who died before reaching one year of age, divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
\item \textbf{Postneonatal Mortality 2005--2008}: Number of children alive for at least 28 days who died before reaching one year of age, divided by the total number of children who lived for 28 days, multiplied by 1,000 (source: DataSUS)
\item \textbf{Enrollment Elementary School 2005--2008}: Number of children enrolled in elementary schools, averaged by classroom size (K--4) (source: INEP).
\item \textbf{Quality of Elementary School 2005--2008}: Average municipal IDEB scores. IDEB scores are estimated as follows. The Ministry of Education takes the average of the students' grades in math and Portuguese, then multiply the those numbers by the student yearly approval rate. INEP reweights the estimators to avoid schools and classroom specific effects (source: INEP, years: 2005, 2007).
\item \textbf{Size of Mayoral Pre-Electoral Coalition 2004}: Number of elected councilors in the mayoral pre-electoral coalition (source: TSE)
\item \textbf{Number of Appointed Bureaucrats 2005--2008}: Number of bureaucrats of the direct administration who were appointed to jobs in the municipality (source: IBGE, years: 2005, 2006, and 2008)
\item \textbf{Number of Females Elected 2004}: Number of females elected to city council (source: TSE)
\item \textbf{Number of Non-Whites Elected 2004}: Number of non-whites elected to the city council. We collect this data for municipalities less than 10,000 inhabitants away from the cutoffs (source: own compilation based on the TSE candidate pictures).
\item \textbf{Competition per Seat 2004}: Number of people running for city councilor divided by the city council size (source: TSE)
\item \textbf{Proportion Approved Legislation 2005}: Number of approved legislation in 2005 divided by the number of proposed legislation (source: InterLegis)
\end{itemize}

@@

\begin{itemize}
\item \textbf{Councilors Job Appointments}: Whether mayors appoint supporters of the city councilors to bureaucratic jobs (source: survey).
\item \textbf{Councilors Demands}: Whether mayors satisfy the public service demands of an allied councilor (e.g., build a health clinic in a councilor's neighborhood) (source: survey).
\item \textbf{Councilors Personal Requests}: Whether mayors satisfy the private requests of an allied councilor (e.g., hire an ally in the bureaucracy) (source: survey).
\item \textbf{Councilors Legislation}: Whether mayors support the councilors' legislative agenda (source: survey).
\item \textbf{Councilors Constructions}: Whether mayors support councilors' requests for more public works (e.g. fix potholes in the councilor's neighborhoods) (source: survey).
\end{itemize}

This Appendix contains other questions we asked in the survey. We collected them to improve our knowledge about the perceptions city councilors have about the roles the municipality and themselves play in the government. We omitted these questions in the main paper because they are not of general interest to our argument. The descriptive statistics of the variables used in the main article follow in Table \ref{tab:descrvarspaper}.

@@

Regarding the first assumption, \citet{McCrary2008} proposes a measure of the distributional imbalance around the discontinuity, testing whether cases are more abundant on either side of the cutoff. In our sample, the McCrary statistic is 0.391 (SE $=$ 0.299), showing no evidence of manipulation.

We run the \citet{cattaneo2019simple} manipulation test, which is based on the density of the local polynomial estimator. We use local polynomials ranging from first to fourth order. As the null hypothesis implies that there is no manipulation, the $p$-values for each polynomial order are: local linear ($p$-value $=0.442$); quadratic ($p$-value $=0.740$); cubic ($p$-value $=0.998$); and quartic ($p$-value $=0.620$). Therefore, we also see no evidence of manipulation.

@@

As Brazil has only a few municipalities with more than one million inhabitants, we focus on cities with less than $571,429 + 47,619 = 619,048$ inhabitants. Table \ref{tab:citycouncilsize2004short} displays the number of municipalities by the thresholds we use in this paper.

@@

To provide an overview of the data dispersion, Figure \ref{sampleplot} plots the municipalities by their proximity to the cutoffs. In the map, we give the contour of every Brazilian city, coloring it by the proximity to the council size thresholds. Darker colors represent municipalities closer to the population thresholds. In regression discontinuity models, the closer a city is to the cutoff, the more influential it is in the estimation. Figure \ref{sampleplot} also shows that the municipalities around the cutoff are reasonably well distributed across the country.

@@

When estimating regression discontinuities with multiple cutoffs, authors usually pool all the discontinuities together to estimate the treatment effect on the outcomes of interest \citep{cattaneo2016interpreting}. Let $C = \{c_0, c_1, \cdots, c_n, c_{n+1}\}$ be the set of cutoffs thresholds associated with the running variable $X$. The outcome variable is assumed to be $Y$. In the pooled model, the Local Average Treatment Effect (LATE) is:

@@

Nevertheless, this may be problematic in our case as municipal characteristics may vary between cutoffs. For instance, population imbalance is a known source of heterogeneity, as shown by \citet{Cattaneo2016}. To ensure that cases are comparable, we estimate the average LATE for all cutoffs. We add a set of controls $Z_i$ to reduce heterogeneity. For instance, the population that determines the change from one threshold to another also determines all the cutoffs. As the population runs smoothly around each cutoff, it makes this variable an ideal control to add to our models. The new estimator is:

@@

\begin{figure}[htb!]
  \begin{center}
    \begin{minipage}{1\linewidth}
      \centerline{\includegraphics[width=5in,height=4in,angle=0]{../figures/bwSens.pdf}}
      \caption{Bandwidth Sensitivity -- Main Models}
      \bigskip
      \label{graphbw}
    \end{minipage}
  \end{center}
\end{figure}

@@

\begin{figure}[H]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[width=5in,height=4in,angle=0]{../figures/placBwSens.pdf}}
      \caption{Bandwidth Sensitivity -- Placebo Regressions}
      \bigskip
      \label{placgraphbw}
    \end{minipage}
  \end{center}
\end{figure}

@@

\begin{figure}[H]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[height=5.9in,width=0.9\linewidth,angle=0]{../figures/dropStateSens1.pdf}}
      \caption{Sensitivity Analysis for the States in the Sample -- Welfare Outcomes}
      \bigskip
      \label{graphsensstatedrop1}
    \end{minipage}
  \end{center}
\end{figure}

\begin{figure}[H]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[height=8in,width=0.9\linewidth,angle=0]{../figures/dropStateSens2.pdf}}
      \caption{Sensitivity Analysis of State Characteristics -- Mechanism Outcomes}
      \bigskip
      \label{graphsensstatedrop2}
    \end{minipage}
  \end{center}
\end{figure}

\begin{figure}[H]
  \begin{center}
      \centerline{\includegraphics[width=0.9\linewidth,angle=0]{../figures/placDropStateSens1.pdf}}
      \caption{Sensitivity Analysis of State Characteristics with Placebo Cutoffs -- Welfare Outcomes}
      \bigskip
      \label{placgraphsensstatedrop1}
  \end{center}
\end{figure}

@@

\begin{figure}[htb!]
  \begin{center}
      \centerline{\includegraphics[width=0.9\linewidth,angle=0]{../figures/placDropStateSens2.pdf}}
      \caption{Sensitivity Analysis of State Characteristics with Placebo Cutoffs -- Mechanism Outcomes}
      \bigskip
      \label{placgraphsensstatedrop2}
  \end{center}
\end{figure}

@@

\begin{figure}[H]
  \begin{center}
    \begin{minipage}{\linewidth}
      \centerline{\includegraphics[height=7in,angle=0]{../figures/ffSens.pdf}}
      \caption{Sensitivity Analysis of Model Functional Form}
      \bigskip
      \label{graphsensff}
    \end{minipage}
  \end{center}
\end{figure}

@@

\begin{figure}[H]
  \begin{center}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/placFfSens.pdf}}
      \caption{Sensitivity Analysis of Model Functional Form -- Placebo Cutoffs}
      \bigskip
      \label{placgraphsensff}
  \end{center}
\end{figure}

@@

## Sensitivity Analysis of Additional Cutoffs

@@

\begin{figure}[H]
\begin{center}
\centerline{\includegraphics[width=6in,height=6.5in,angle=0]{../figures/rdAddCuts1.pdf}}
\caption{Sensitivity to Additional Cutoffs -- Welfare Outcomes}
\bigskip
\label{senscut1}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\centerline{\includegraphics[width=6in,angle=0]{../figures/rdAddCuts2.pdf}}
\caption{Sensitivity to Additional Cutoffs -- Mechanism Outcomes}
\bigskip
\label{senscut2}
\end{center}
\end{figure}
